Verifying Interactive WebSocket Applications in TypeScript

### 2020-01-23 PLACES_2020/main.tex
change

\title{Multiparty Session Type-safe Web Development in TypeScript}
\author{Anson Miu
\institute{Imperial College London}
\and
Francisco Ferreira
\institute{Imperial College London}
\and
Nobuko Yoshida
\institute{Imperial College London}
\and
Fangyi Zhou
\institute{Imperial College London}
}
\def\titlerunning{MPST-safe Web Development in TypeScript}
\def\authorrunning{A. Miu, F. Ferreira, N. Yoshida \& F. Zhou}
\begin{document}
\maketitle

\begin{abstract}
Advancements in mobile device compute power has made interactive web applications possible, allowing the web browser to render content dynamically and support low-latency communication with the server. This comes at a cost to the developer, who now need to reason more about the communication correctness of their application as they support more complex communication patterns.

@@

We see the application of the MPST methodology to verifying interactive web applications written in TypeScript to be an interesting design space - to what extent can the MPST methodology be applied such that we leverage TypeScript language features to maximise static safety guarantees whilst preserving a flexible, natural and idiomatic workflow for TypeScript developers? Such a workflow would ultimately decrease the overhead for incorporating MPST into mainstream web development, which reduces development time by programmatically verifying implementations for communication correctness.

@@

We leverage the Scribble toolchain to check for protocol well-formedness. This directly corresponds to the multiparty session type theory of verifying valid local type \textit{projections} for all participants of a global type. The implementation of this validation by the toolchain follow from the algorithmic projection procedures defined in \cite{MPST}. We obtain \textit{endpoint protocols} for each role in a well-formed global protocol. An endpoint protocol only preserves the interactions defined by the global protocol in which the target role is involved, and corresponds to an equivalent \textit{Endpoint Finite State Machine} (EFSM) in the Scribble toolchain. We use the EFSMs as a basis for API generation and adopt the formalisms in \cite{Hybrid2016}.

@@

Developers can validate the communication aspects of their implementation against their EFSM to verify protocol conformance. Our approach integrates the EFSM into the development workflow by encoding session types as TypeScript types. We appreciate the inherent differences in development considerations between front- and back-end web development: business logic tends to initiate communication in the back-end, whilst communication in the front-end is mostly driven by user interactions with the browser. This motivates our design choice of generating different session types encodings for the target application on the server (\sectionref{section:server}) and in the browser (\sectionref{section:browser}).

@@

We will refer to the \texttt{Svr} EFSM (fig.~\ref{fig:efsmsvr}) as a running example in this section. For server-side targets, we encode EFSM states as TypeScript types depending on the type of state. We adopt the EFSM definition presented in \cite{Hybrid2016} to consider receive and send states separately. We assign each TypeScript encoding its state identifier from the EFSM as its type alias, thus providing syntactic sugar when referring to the successor state in the TypeScript encoding of the current state. For any state $S$ in the EFSM, we refer to the TypeScript type alias of its encoding as $\llbracket S \rrbracket$.

We make a key design decision \textit{not} to expose communication channels in the TypeScript session type encodings to provide linearity guarantees (\sectionref{section:serverlinear}). Our encodings sufficiently exposes seams for the developer to inject their business logic, whilst the generated session API (\sectionref{section:serversessionapi}) handles the sending and receiving of messages; as a result, the encodings do not concern with the role involved in the send/receive action. We outline the encodings below using examples from the \textit{Noughts and Crosses} game server (Listing \ref{lst:svr}).

@@

We will refer to the \texttt{P1} EFSM (figure \ref{fig:efsmp1}) as a running example in this section. Preserving behavioural typing and channel linearity are known to be challenging for browser-side applications due to the event-driven nature of user interaction: in the case of \textit{Noughts and Crosses}, once the user makes a move by clicking on a cell on the game board, this click event must be deactivated until the user's next turn, otherwise the user can click again and violate channel linearity. Our design goal is to enforce this statically through the APIs we generate.

@@

\paragraph{Components} A component is a reusable UI element which contains its own markup and logic. Components implement a \texttt{render()} function which returns a React element, the smallest building blocks of a React application. Components can keep \textit{state} and the \texttt{render()} function is invoked upon a change of state. A simple counter can be implemented as a component, with the count stored as state, a button which increments the count when clicked and a \texttt{div} that renders the current count. Components can also render other components, which gives rise to a parent/child relationship between components. Parents can pass data to children as \textit{props}: going back to the aforementioned example, the the counter component could render a child component \texttt{<StyledDiv count=\{this.state.count\} />} in its \texttt{render()} function, propagating the \texttt{count} from its state to the child. This enables reusability, and for our use case, gives control to the parent on what data to pass to its children (e.g. pass the payload of a received message to a child to render).

>>>>

\title{Verifying Interactive WebSocket Applications in TypeScript}
\author{Anson Miu
\institute{Imperial College London}
\and
Francisco Ferreira
\institute{Imperial College London}
\and
Nobuko Yoshida
\institute{Imperial College London}
\and
Fangyi Zhou
\institute{Imperial College London}
}
\def\titlerunning{Verifying Interactive WebSocket Applications in TypeScript}
\def\authorrunning{A. Miu, F. Ferreira, N. Yoshida \& F. Zhou}
\begin{document}
\maketitle

\begin{abstract}
Advancements in mobile device compute power has made interactive web applications possible, allowing the web browser to render content dynamically and support low-latency communication with the server. This comes at a cost to the developer, who now needs to reason more about the communication correctness of their application as they support more complex communication patterns.

@@

We see the application of the MPST methodology to verifying
interactive web applications written in TypeScript to be an
interesting design space -- to what extent can the MPST methodology be
applied such that we leverage TypeScript language features to maximise
static safety guarantees whilst preserving a flexible, natural and
idiomatic workflow for TypeScript developers? Such a workflow would
ultimately decrease the overhead for incorporating MPST into
mainstream web system, which reduces development time by
programmatically
verifying communication correctness of the implementation.

@@

We leverage the Scribble toolchain to check for protocol
well-formedness. This directly corresponds to the multiparty session
type theory of verifying valid local type \textit{projections} for all
participants of a global type. The implementation of this validation
by the toolchain follows from the algorithmic projection procedures
defined in \cite{MPST}. We obtain a set of \textit{endpoint protocols}
-- one for each role from a well-formed global protocol. An endpoint protocol only preserves the interactions defined by the global protocol in which the target role is involved, and corresponds to an equivalent \textit{Endpoint Finite State Machine} (EFSM) in the Scribble toolchain. We use the EFSMs as a basis for API generation and adopt the formalisms in \cite{Hybrid2016}.

@@

Developers can validate the communication aspects of their
implementation against their EFSM to verify protocol conformance. Our
approach integrates the EFSM into the development workflow by encoding
session types as TypeScript types.
There are inherent differences between front- and back-end web development: business logic tends to initiate communication in the back-end, whilst communication in the front-end is mostly driven by user interactions with the browser. This motivates our design choice of generating different session types encodings for the target application on the server (\sectionref{section:server}) and in the browser (\sectionref{section:browser}).

@@

We will refer to the \texttt{Svr} EFSM (fig.~\ref{fig:efsmsvr}) as a
running example in this section. For server-side targets, we encode
EFSM states as TypeScript types depending on the type of state. We
adopt the EFSM definition presented in \cite{Hybrid2016} to consider
receive and send states separately. Each TypeScript encoding is
assigned to its state identifier from the EFSM as its type alias, thus providing syntactic sugar when referring to the successor state in the TypeScript encoding of the current state. For any state $S$ in the EFSM, we refer to the TypeScript type alias of its encoding as $\llbracket S \rrbracket$.

We make a key design decision \textit{not} to expose communication channels in the TypeScript session type encoding to provide linearity guarantees (\sectionref{section:serverlinear}). Our encoding sufficiently exposes seams for the developer to inject their business logic, whilst the generated session API (\sectionref{section:serversessionapi}) handles the sending and receiving of messages; as a result, the encoding does not concern with the role involved in the send/receive action. We outline the encoding below using examples from the \textit{Noughts and Crosses} game server (Listing \ref{lst:svr}).

@@

We will refer to the \texttt{P1} EFSM (figure \ref{fig:efsmp1}) as a
running example in this section. Preserving behavioural typing and
channel linearity are known to be challenging for browser-side
applications due to the event-driven nature of user interaction: in
the case of \textit{Noughts and Crosses}, once the user makes a move
by clicking on a cell on the game board, this click event must be
deactivated until the user's next turn, otherwise the user can click
again and violate channel linearity. Our design goal is to enforce
this statically through the generated APIs.

@@

\paragraph{Components} A component is a reusable UI element which
contains its own markup and logic. Components implement a
\texttt{render()} function which returns a React element, the smallest
building blocks of a React application. Components can keep
\textit{state} and the \texttt{render()} function is invoked upon a
change of state. A simple counter can be implemented as a component,
with the count stored as state, a button which increments the count
when clicked and a \texttt{div} that renders the current
count. Components can also render other components, which gives rise
to a parent/child relationship between components. Parents can pass
data to children as \textit{props}: going back to the aforementioned
example, the the counter component could render a child component
\texttt{<StyledDiv count=\{this.state.count\} />} in its
\texttt{render()} function, propagating the \texttt{count} from its
state to the child. This enables reusability, and for our use case,
gives a control to the parent on what data to pass to its children (e.g. pass the payload of a received message to a child to render).
